---
title: "Project proposal"
author: "Adler and Jackie"
output: github_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(readxl)
```

## 1. Introduction

The data is of the presence of birds around an aquaculture farm, to see if the aquaculture farm impacted the population of birds in the area, and if those bird populations were negatively impacted by the planes coming from the nearby airport. It was collected by John Anderson and a group of research assistants. They went out a few days each month, and for an hour they collected data on the environmental factors (including temperature, wind), the tide, and what species were observed.

We are looking to explore any patterns in the presence of bird species, and which birds are found in specific whether conditions. 


## 2. Data

```{r load_data}

seabird_count <- read_excel("../data/Copy of Trenton seabird count_nov 2014-2023.xls")

seabird_count <- seabird_count %>%
  rename(observer = "OBSERVER",
         date = "DATE",
         hours = "HOURS",
         time = "TIME",
         temp = "TEMP (F)",
         wind_speed = "WIND SPEED (mph)",
         wind_direction = "WIND DIRECTION",
         tide_obs = "TIDE OBS",
         tide_percentage = "TIDE %",
         weather_obs = "WEATHER OBS",
         weather_percentage = "WEATHER %",
         precipitation = "PRECIPITATION",
         total = "Total",
         herring_gull = "Herring Gull",
         laughing_gull = "Laughing Gull",
         great_black_backed_gull = "Great Black-backed Gull",
         red_throated_loon = "red throated loon",
         common_loon = "Common Loon",
         long_tailed_duck = "Long-tailed Duck",
         red_breasted_merganser = "Red breasted Merganser",
         common_merganser = "Common Merganser",
         black_duck = "Black Duck",
         surf_scoter = "surf scoter",
         common_eider = "Common Eider",
         bufflehead = "Bufflehead",
         goldeneye = "Goldeneye",
         lesser_scaup = "Lesser Scaup",
         black_scoter = "Black Scoter",
         white_winged_scoter = "White-Winged Scoter",
         red_necked_grebe = "Red-Necked Grebe",
         horned_grebe = "Horned Grebe",
         bonapartes_gull = "Bonaparte's Gull",
         belted_kingfisher = "Belted Kingfisher",
         common_tern = "Common Tern",
         ring_billed_gull = "Ring-Billed Gull",
         canada_goose = "Canada Goose", 
         great_blue_heron = "Great Blue Heron",
         double_crested_cormorant = "Double-Crested Cormorant",
         common_crow = "Common Crow",
         black_guillemot = "Black Guillemot",
         osprey = "Osprey",
         mallard_duck = "Mallard Duck",
         spotted_sandpiper = "Spotted Sandpiper",
         semipalmated_plover = "Semipalmated Plover",
         bald_eagle = "Bald Eagle",
         red_tailed_hawk = "Red-Tailed hawk",
         turkey_vulture = "Turkey Vulture",
         rough_legged_hawk = "rough legged hawk",
         harrier = "Harrier",
         kestrel = "Kestrel",
         turkey = "Turkey",
         unidentifiable_duck = "Unidentifiable Duck",
         unidentifiable = "Unidentifiable (too far to ID)")

```

```{r remove_columns_and_rows}
seabird_count <- seabird_count %>%
 select(-"...17", -"COMMENTS", -"...59", -"...60") %>%
  slice(-1) %>%
mutate(across(total:unidentifiable, as.numeric)) %>%
  pivot_longer(
    cols = total:unidentifiable,
      names_to = "species",
   values_to = "count")
  
```
```{r tidy_data}
seabird_count_tidy <- seabird_count  %>%
  mutate(count = str_replace_na(count, "0")) %>%
  mutate(count = as.double(count)) %>%
  mutate(species = str_replace(species, "Surf Scoter", "surf_scoter")) %>%
  mutate(tide_obs = str_replace(tide_obs, "half", "mid")) %>%
  mutate(tide_obs = str_replace(tide_obs, "higgh", "high")) %>% 
  mutate(tide_obs = str_replace(tide_obs, "hightide", "high")) %>% 
  mutate(tide_obs = str_replace(tide_obs, "mid-high", "high")) %>%
  mutate(tide_obs = str_replace(tide_obs, "thre quarters", "high")) %>%
  mutate(tide_obs = str_replace(tide_obs, "three quarters", "high")) %>%
  mutate(tide_obs = str_replace(tide_obs, "threequarters", "high")) %>%
  mutate(tide_obs = str_replace(tide_obs, "med", "mid")) %>%
  mutate(tide_obs = str_replace(tide_obs, "mid-low", "low")) %>%
  mutate(tide_obs = str_replace(tide_obs, "mlow", "low")) %>%
  mutate(tide_obs = str_replace(tide_obs, "one third", "low")) %>%
  mutate(tide_obs = str_replace(tide_obs, "quarter", "low")) %>%
  mutate(tide_obs = str_replace(tide_obs, "third", "low")) 

seabird_count_tidy <- seabird_count %>%
  mutate(weather_obs = str_replace(weather_obs, "clear, runny", "clear")) %>%
  mutate(weather_obs = str_replace(weather_obs, "clear, sunny", "clear")) %>%
  mutate(weather_obs = str_replace(weather_obs, "sunny, whitecaps present", "clear")) %>%
  mutate(weather_obs = str_replace(weather_obs, "sunny, breezy", "clear")) %>%
  mutate(weather_obs = str_replace(weather_obs, "sunny!", "clear")) %>%
  mutate(weather_obs = str_replace(weather_obs, "few clouds", "clear")) %>%
  mutate(weather_obs = str_replace(weather_obs, "mosly clear", "clear")) %>%
  mutate(weather_obs = str_replace(weather_obs, "mostly sunny", "clear")) %>%
  mutate(weather_obs = str_replace(weather_obs, "clody, breezy, waves 1-2 feet", "cloudy")) %>%
  mutate(weather_obs = str_replace(weather_obs, "cloud", "cloudy")) %>%
  mutate(weather_obs = str_replace(weather_obs, "clouds", "cloudy")) %>%
  mutate(weather_obs = str_replace(weather_obs, "mostly cloudy", "cloudy")) %>%
  mutate(weather_obs = str_replace(weather_obs, "partly coudy", "cloudy")) %>%
  mutate(weather_obs = str_replace(weather_obs, "cloufy", "cloudy")) %>%
  mutate(weather_obs = str_replace(weather_obs, "high cloudy", "cloudy")) %>%
  mutate(weather_obs = str_replace(weather_obs, "hih clouds", "cloudy")) %>%
  mutate(weather_obs = str_replace(weather_obs, "foggy, low visability", "fog")) %>%
  mutate(weather_obs = str_replace(weather_obs, "hazy", "fog")) %>%
  mutate(weather_obs = str_replace(weather_obs, "high haze", "fog")) %>%
  mutate(weather_obs = str_replace(weather_obs, "high overcast", "overcast")) %>%
  mutate(weather_obs = str_replace(weather_obs, "low overcast", "overcast")) %>%
  mutate(weather_obs = str_replace(weather_obs, "cloudy w drizzle", "rain")) %>%
  mutate(weather_obs = str_replace(weather_obs, "cloudy / rain", "rain")) %>% 
  mutate(weather_obs = str_replace(weather_obs, "cloudyy", "cloudy")) %>% 
  mutate(weather_obs = str_replace(weather_obs, "partly cloudyy", "cloudy")) %>% 
  mutate(weather_obs = str_replace(weather_obs, "sunny,", "sunny")) %>% 
  mutate(weather_obs = str_replace_all(weather_obs, "sunny, white caps present", "sunny")) %>% 
  mutate(weather_obs = str_replace(weather_obs, "clear!", "clear")) %>% 
  mutate(weather_obs = str_replace(weather_obs, "rainy", "rain")) %>% 
  mutate(weather_obs = str_replace(weather_obs, "mostly clear", "clear")) %>% 
  mutate(weather_obs = str_replace(weather_obs, "cloudys", "cloudy")) %>% 
  mutate(weather_obs = str_replace(weather_obs, "cloudyy, breezy, waves 1-2 feet", "cloudy")) %>% 
  mutate(weather_obs = str_replace(weather_obs, "cloudyy/rain", "cloudy")) %>% 
  mutate(weather_obs = str_replace(weather_obs, "hih cloudy", "cloudy")) %>% 
  mutate(weather_obs = str_replace(weather_obs, "foggy", "fog")) %>% 
  mutate(weather_obs = str_replace(weather_obs, "cloudyy with drizzle	", "rain")) %>%
  mutate(weather_obs = str_replace(weather_obs, "cloudyy", "cloudy")) %>% 
  mutate(weather_obs = str_replace(weather_obs, "cloudy with drizzle", "cloudy")) %>% 
  mutate(weather_obs = str_replace(weather_obs, "sunny white caps present", "sunny"))

```

```{r glimpse}

glimpse(seabird_count)
unique(seabird_count$species)
unique(seabird_count$count)
unique(seabird_count$weather_obs)

seabird_count_tidy %>%
  # mutate(weather_obs = str_replace_all(weather_obs, "sunny white caps present", "sunny")) %>%
  # mutate(weather_obs = str_replace(weather_obs, "cloudyy", "cloudy")) %>%
  distinct(weather_obs)
```

## 3. Ethics review

This research does not include people, so we are not concerned with an ethics review.

## 4. Data analysis plan

The variables we are visualizing are the species count, season, temperature, wind speed, and tide observation. We will be looking to see if the variables for a certain group of birds have similarities or differences to the variables for other groups of birds. 

We won't need any additional data, but we will need to do a little research on bird groupings (for example, ducks vs gulls). 

```{r preliminary_analysis_month_count}

ggplot(seabird_count, aes(x = month, fill = count)) +
  geom_bar() +
  labs(x = "Month", y = "Number of Birds", title = "Number of Birds Recorded Each Month", subtitle = "2014 - 2024")

```

We think that bar graphs and scatter plots will be most useful in our data visualization. 
